---
title: "Missouri River Basin Rasters"
author: "Mandy Liesch"
date: "8/16/2022"
output:
  html_document:
    toc: true
    code_folding: hide

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

## Introduction

The objective of this webpage is to visualize and review the underlying raster data that informs the input data calibrated from the Missouri River basin ORISE project on GCAM with the EPA. There are 5 different soil carbon rasters that are utilized in the calculation of land use based emissions for both soil and vegetative carbon. One basin is focused on, as it has the largest percent growth in the United States between 2020 and 2060, with a 5%+ growth in cropland with corresponding losses in unmanaged pasture. The five soil carbon rasters represent estimations in model based soil carbon. 


```{r}
#Load the required packages
library(sp)
library(sf)
library(terra)
library(tidyverse)
library(tidyterra)
library(raster)
library(rasterVis)
library(ggplot2)
library(leaflet)
library(RColorBrewer)
library(ncdf4) 
```

## Materials and Methods

There are several different rasters in this Missouri River Basin Analysis: 

### Soil Carbon Rasters: 30 cm

- GCAM 30 cm estimate:  Original GCAM Estimates:  (recreated with the inital .csv file and the land use in house raster). The ratio of the 30 cm/100 cm HWSD raster was used to multiply with the 100 cm GCAM raster to calculate the approximate level of soil carbon at 30 cm level (for comparison)  
- FAO Glosis 2018 Raster: From the [FAO Data](https://www.fao.org/soils-portal/data-hub/soil-maps-and-databases/global-soil-organic-carbon-map-gsocmap/en/), and is a fusion of many different sytles, with in country data, filled in with soilgrids 2017 for countries that provided no data.
- Harmonized World Soils Database: The topsoil (30 cm) raster from the [Regridded Harmonized World Soil Database v1.2](https://daac.ornl.gov/SOILS/guides/HWSD.html).    
- SoilGrids 2017 30 cm raster: Derived with random forest machine learning algorithms [Hengel et al., 2017](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0169748), downloaded the 30 cm soil carbon stocks 250 m resoultion.   
- SoilGrids 2020 30 cm raster: downloaded using the [WebDav Protocol](https://www.isric.org/explore/soilgrids/soilgrids-access)  

### Vegetation Rasters

- Initial GCAM Vegetation Estimates from the same Houghton Data, with the initial raster used to create the current vegetation estimates


### Land Use and Change Rasters

- In House Land Use Raster, based on the European Space Agency (ESA) and Ramunkutty Tundra and Pasture Estimates. 300 m x 300 m resolution.   
- Historic Land Use from 1700 from [Li et al., 2023](https://essd.copernicus.org/articles/15/1005/2023/)
- Future Land Use compilation (2060) from [Chen et al,. 2020](https://www.nature.com/articles/s41597-020-00669-x) mean model RCP4.5 SSP2 and the [Land Use Harmonization v2](https://luh.umd.edu/index.shtml) RCP4.5 SSP2 described in [Hurtt et al., 2020](https://gmd.copernicus.org/articles/13/5425/2020/gmd-13-5425-2020-discussion.html) for the pasture and rangeland land use categories. Categories were totaled into the GCAM land use designations and filtered by most prominent. 
- Cropland Land Use Change from 2000-2019, derived from [WRI](https://glad.umd.edu/dataset/croplands) Data. 3 km resolution, and the value is the % of land converted to cropland in that time period.  


```{r, results=FALSE}
MissouriSV<-vect("missouri_river_basin.shp")


Miss_2020<-rast("Rasters/Global_LU_2020.tif")
Miss_LUC_2020<-project(Miss_2020, 'EPSG:4326',  method = "near")
levels(Miss_LUC_2020) <- data.frame(id=c(1,2,3,4, 5, 6, 7), cover=c("Cropland", "Forest", "Grassland",  "Shrubland", "Urban","Rock/Desert", "Pasture"))
Missouri_2020<-crop(Miss_LUC_2020, MissouriSV)
Miss_LULC<-mask(Missouri_2020, MissouriSV) 

#Load in the In House Categorical Land Use Raster
Missouri_LUC<-rast("Rasters/Missouri_LUC.tif")
Missouri_LUC_84<-project(Missouri_LUC, 'EPSG:4326',  method = "near")
levels(Missouri_LUC_84) <- data.frame(id=c(1,2,3,4, 5, 6, 8), cover=c("Cropland", "Forest", "Grassland",  "Shrubland", "Urban","Rock/Desert", "Pasture"))

Hist_LUC<-rast("Rasters/conus_lulc_1700.tif")
Hist_LULC<-project(Hist_LUC, 'EPSG:4326',  method = "near")
Hist_LULC <- ifel(Hist_LULC == 0, 6, Hist_LULC)

Missouri_Hist<-crop(Hist_LULC, MissouriSV)
Miss_Hist_LULC<-mask(Missouri_Hist, MissouriSV) 

levels(Miss_Hist_LULC)<-data.frame(id=c(4, 5, 6, 7, 8, 9), cover=c("Forest", "Shrubland", "Grassland", "Wetland", "Water", "Rock/Desert"))

Miss_Future<-rast("Rasters/Global_Future_LU_2070.tif")
Missouri_fut<-crop(Miss_Future, MissouriSV)
Miss_Fut_LULC<-mask(Missouri_fut, MissouriSV)

levels(Miss_Fut_LULC) <- data.frame(id=c(1,2,3,4, 5, 7), cover=c("Cropland", "Forest", "Grassland",  "Shrubland", "Urban", "Pasture"))


#Load in the WRI Land Use Cropland Change
Missouri_crop_gain<-rast("Rasters/Missouri_CropGain.tif")
Missouri_crop_gain_84<-project(Missouri_crop_gain, 'EPSG:4326')


#All 30 cm Rasters
FAO<-rast("Rasters/Missouri_FAO.tif")
Missouri_FAO<-project(FAO, 'EPSG:4326')
Missouri_FAO<-classify(Missouri_FAO, cbind(150, Inf, 150))

GCAM<-rast("Rasters/Missouri_GCAM.tif")
Missouri_GCAM<-project(GCAM, 'EPSG:4326')
Missouri_GCAM<-ifel(Missouri_GCAM > 97, 150, Missouri_GCAM)

HWSD<-rast("Rasters/Missouri_HWSD.tif")
Missouri_HWSD<-project(HWSD, 'EPSG:4326')
Missouri_HWSD<-ifel(Missouri_HWSD > 114, 150, Missouri_HWSD)

SG17<-rast("Rasters/Missouri_SG17.tif")
Missouri_SG17<-project(SG17, 'EPSG:4326')
Missouri_SG17<-classify(Missouri_SG17, cbind(150, Inf, 150))

SG20<-rast("Rasters/Missouri_SG20.tif")
Missouri_SG20<-project(SG20, 'EPSG:4326')
Missouri_SG20<-classify(Missouri_SG20, cbind(150, Inf, 150))




```


## Results

### Soil Carbon Rasters

#### 

There are five different soil rasters that are important to this study presented. The intial GCAM dataset relied on a 100 cm version of Houghton, which this study standardized to a 30 cm depth to match the other rasters. The GCAM raster also utilized the 2015 land use map (below) to recreate a visual representation to match the other rasters.

```{r}
SOC30cm_Stack<-c( Missouri_FAO, Missouri_GCAM, Missouri_HWSD, Missouri_SG17, Missouri_SG20) 

names(SOC30cm_Stack) <- c(  "FAO Glosis 30 cm", "Houghton 30 cm", "HWSD 30 cm", "SoilGrids 2017 30 cm", "SoilGrids 2020 30 cm")

#https://rdrr.io/cran/terra/man/plet.html
plet(SOC30cm_Stack, 1:5, main="Soil Carbon\n Content\n (Mha)", tiles="Streets", share=TRUE, collapse=FALSE) |> lines(MissouriSV, lwd=2)

which_min<-which.min(SOC30cm_Stack)
freq_min <- freq(which_min)
freq_min$area <- freq_min$count * 0.0009
freq_min$percent = round(100 * freq_min$area / sum(freq_min$area), 1)

which_max<-which.max(SOC30cm_Stack)
freq_max <- freq(which_max)
freq_max$area <- freq_max$count * 0.0009
freq_max$percent = round(100 * freq_max$area / sum(freq_max$area), 1)

range<-range(SOC30cm_Stack)
```

Half of the minimum values in the raster came from the FAO Glosis raster (~52%), with a majority of this area located in the western portion of the basin (an area compromizing the shrubland and short grass prairies converted into pasture). The HWSD was the minimum value in ~20% of the basin, in the area that mostly comprises the cropland area. The SoilGrids 2020 was the minimum raster value in ~18% of the basin, in the southern portion of the basin, as well as the Canadian portion of the basin. The SoilGrids 2017 values did not have any minimum raster values.

The SoilGrids 2017 has ~58% of the maximum values in each grid cell, with almost all areas represented in the basin except the same southern portion of the basin that the SoilGrids 2020 was the maximum values (Kansas/Nebraska), which had the Houghton values as the maximum values for ~25% of the basin. The SoilGrids 2017 map also has the highest values in the western Woodlands (Bighorn National Forest, Custer, and Yellowstone).

### Missouri River Land Use

#### Historic 1700 Land Use Raster

The Missouri River basin was originally a vast grassland (approx 65%) in the eastern and central portions, with a few forested regions (~20%) in Missouri, the Black Hills, Bighorn National Forest, along the Colorado frontrange, and Yellowstone. The climate gradient gets substantially wetter the further east you head in the basin, with shrubland (~15%) replacing short grass prairie as water becomes more scarce in Montana and Wyoming. This land use changed vastly starting around 1830 in the far east, moving westward at an almost explosive pace. 

```{r}

plet(Miss_Hist_LULC, main="Land Use\n Category, 1700", col=c( "#075149FF", "#F05C3BFF", "#46732EFF", "#192EC0FF", "#192EC0FF", "#8A9197FF"), tiles="Streets") |> lines(MissouriSV, lwd=2)

freq_hist <- freq(Miss_Hist_LULC)
freq_hist$area <- freq_hist$count * 0.0009
freq_hist$percent = round(100 * freq_hist$area / sum(freq_hist$area), 1)

```

#### Present (2015) In House Derived Raster 

The pastureland area in this raster came from a modification of the Ramenkutty definition superimposed on the European land use map. Cropland has taken over a majority of the eastern long grass prairies (~19%), and pasture/rangeland makes up the largest land use type in the basin (~36%), replacing the Shrubland (~5%) and Grassland (~33%) categoires. The remaining forest lands (~8%) are mostly the national forest areas. 

```{r}
#https://rdrr.io/cran/terra/man/factors.html
plet(Missouri_LUC_84, main="Land Use\n Category, 2015", col=c("#FED439FF", "#075149FF", "#46732EFF",  "#F05C3BFF", "#C80813FF", "#8A9197FF", "#FD8CC1FF"), tiles="Streets") |> lines(MissouriSV, lwd=2)

freq_pres <- freq(Missouri_LUC_84)
freq_pres$area <- freq_pres$count * 0.0009
freq_pres$percent = round(100 * freq_pres$area / sum(freq_pres$area), 1)

```

This raster is actually different than the baseline values that the GCAM model utilizes in 2015, quite substantially. The GCAM model has cropland accounting for ~28% of the land area in this basin, with pastureland around ~51%. Grassland accounts for 10% of the basin, and forests just over 6%, and Shrubland around 4%. Knowing these area discrepancies will make the future composite raster make more sense. 

#### Future (2070) Composite Derived Raster

In the future, the agricultural extensification looks more prominent, as the cell size is bigger (0.05 degrees), compared to the 30 m resolution in the 2015 map. There are some misclassification areas, with urban regions showing up in areas that are currently national forests, and places like Kansas City and Omaha not registering as urban, but rather grassland. This map simulation has Cropland nearly doubling (~31%), as well as pastureland (~59%). This could be that some of the shrubland in Wyoming/Montana ARE used for grazing, despite not technically being grasslands. This does NOT match the patterns in the GCAM simulation, which shows a decrease in pastureland at the expense of cropland, but I think this map represents the GCAM land use. Forest remains fairly steady at around 6% of the land basin. 

```{r}
plet(Miss_Fut_LULC, main="Land Use\n Category, 2070", col=c("#FED439FF", "#075149FF", "#46732EFF",  "#F05C3BFF", "#C80813FF", "#FD8CC1FF"), tiles="Streets") |> lines(MissouriSV, lwd=2)

freq_fut <- freq(Miss_Fut_LULC)
freq_fut$area <- freq_fut$count * 0.0009
freq_fut$percent = round(100 * freq_fut$area / sum(freq_fut$area), 1)
```

Compared to the numerical values provided for the GCAM land area, Cropland makes up ~33% of the basin, and pastureland is ~46% of the basin. Grassland and shrubland were underrepresented in this picture (~10% and ~4% respectively), and forests were pretty static (~6%). This is one of the discontinuities from having area additive representations vs the rasterized grid approximations. The GCAM scenario also doesn't represent the SSP2 RCP 4.5 presented above, but rather is its own unique representation of data, which can also be responsible for these discrepancies.

### Cropland Gains

Just for reference, the WRI percent cropland growth between 2000 and 2019 was added here, showing the expansion of cropland areas in some of these grid cells by over 50% in Missouri, along the eastern edge of the basin in the fertile lands, and in Montana. These are also areas that are experiencing increases into the future. 

```{r}


plet(Missouri_crop_gain_84, main="Percent\n Cropland\n Added\n 2000-2019", col=map.pal("reds"), tiles="Streets") |> lines(MissouriSV, lwd=2)

```






